---
title: "p8105_hw2_sc5352.Rmd"
author: "Chrissy Chen"
date: "2023-09-30"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Problem 1  
```{r, message=FALSE}
library(tidyverse)
```

#### clean the data in pols-month.csv
```{r, message=FALSE}
pols_month = 
  read_csv("data/pols-month.csv") |>
  separate(mon, into = c("year", "month", "day"), sep = "-") |>
  mutate(
    year = as.integer(year), 
    month = as.integer(month), 
    day = as.integer(day),
    month = month.abb[month],
    president = ifelse(prez_gop == 1, "gop", "dem")
    ) |>
  select(year, month, everything(), -prez_gop, -prez_dem, -day)
```

#### clean the data in snp.csv 
```{r, message=FALSE}
snp = 
  read_csv("data/snp.csv") |>
  separate(date, c("month", "day", "year"), "/") |>
  mutate(
    month = as.integer(month),
    month = month.abb[month],
    year = as.integer(year),
    year = if_else(year > 2023, year - 100, year),
    day = as.integer(day)
  ) |>
  select(year, month, everything())
```

#### tidy the unemployment data
```{r, message=FALSE}
unemployment = 
  read_csv("data/unemployment.csv") |>
  rename(year = Year) |>
  pivot_longer(
    Jan:Dec, 
    names_to = "month",
    values_to = "unemployment"
  ) |>
  select(year, month, unemployment)
```
```{r, message=FALSE}
data_538 = 
  left_join(pols_month, snp) |>
  left_join(x = _, y = unemployment)
```


### Problem 2

```{r, message=FALSE}
library(dbplyr)
library(readxl)
library(tidyverse)
```

```{r, message=FALSE}
Mr_Trash_Wheel = 
  read_excel("Trash Wheel.xlsx", sheet = "Mr. Trash Wheel") |>
  janitor::clean_names() |> 
  separate(date, into = c("year", "mon", "day"), sep = "-") |>
  mutate(
    homes_powered = (weight_tons*500)/30,
    type = "Mr_Trash_Wheel"
  ) |>
  drop_na(dumpster)|>
  select(-mon,-x15,-x16) 

Prof_Trash_Wheel = 
  read_excel("Trash Wheel.xlsx", sheet = "Professor Trash Wheel") |>
  janitor::clean_names() |>
  separate(date, into = c("year", "mon", "day"), sep = "-") |>
  mutate(
    homes_powered = (weight_tons*500)/30,
    type = "Prof_Trash_Wheel",
    year = as.character(year), 
    month = as.character(month)
  )|>
  drop_na(dumpster)|>
  select(-mon)

Gwynnda = 
  read_excel("Trash Wheel.xlsx", sheet = "Gwynnda Trash Wheel") |>
  janitor::clean_names() |>
  separate(date, into = c("year", "mon", "day"), sep = "-") |>
  mutate(
    homes_powered = (weight_tons*500)/30,
    type = "Gwynnda",
    year = as.character(year), 
    month = as.character(month)
  ) |>
  drop_na(dumpster)|>
  select(-mon)

#combine_data = bind_rows(Mr_Trash_Wheel,Prof_Trash_Wheel)
all_data = bind_rows(Mr_Trash_Wheel,Prof_Trash_Wheel,Gwynnda)
```

The dataset `Mr_Trash_Wheel` has `r nrow(Mr_Trash_Wheel)` observations and `r ncol(Mr_Trash_Wheel)` variables. The `homes_powered` column has been added based on the calculations. Below shows a part of the dataset `Mr_Trash_Wheel` Variables are `r names(Mr_Trash_Wheel)`

```{r}
head(Mr_Trash_Wheel)
```

The dataset `Prof_Trash_Wheel` has `r nrow(Prof_Trash_Wheel)` observations and `r ncol(Prof_Trash_Wheel)` variables. The total weight of trash collected by Professor Trash Wheel is `r sum(Prof_Trash_Wheel$weight_tons)`. 
The dataset `Gwynnda` has `r nrow(Gwynnda)` observations and `r ncol(Gwynnda)` variables. The total number of cigarette butts collected by Gwynnda in July of 2021 `r sum(filter(Gwynnda, month == "July", year == "2021")$cigarette_butts)`. 
The combined dataset `all_data` has `r nrow(all_data)` observations and `r ncol(all_data)` variables.

### Problem 3
```{r, message=FALSE}
baseline = 
  read_csv("data_mci/MCI_baseline.csv", skip = 1) |>
  janitor::clean_names() |>
  mutate(
    sex = ifelse(sex == 1, "male", "female"),
    apoe4 = ifelse(apoe4 == 1, "carrier", "noncarrier"))

baseline_mci = 
  read_csv("data_mci/MCI_baseline.csv", skip = 1) |>
  janitor::clean_names() |>
  mutate(
    sex = ifelse(sex == 1, "male", "female"),
    apoe4 = ifelse(apoe4 == 1, "carrier", "noncarrier"),
    age_at_onset = ifelse(age_at_onset == ".", NA, age_at_onset)
    ) |>
  filter(current_age < age_at_onset | is.na(age_at_onset))

```

When importing the `baseline` and `amyloid` datasets, I deleted the first line so that the second line became the headers. Also, `sex` and `apoe4` variables have been recoded based on the instructions. 

In the dataset `baseline_mci`, I removed participants who did not meet the stated inclusion criteria: changing "." into NA and filter out the target rows with conditions that current age is less than age_at_onset or the age_at_onset is NA.  

There are `r nrow(baseline)` participants, and `r nrow(baseline_mci)` of those develop MCI. The average baseline age is `r mean(baseline_mci$current_age)`. `r nrow(filter(baseline_mci, sex == "female", apoe4 == "carrier")) / nrow(filter(baseline_mci, sex == "female")) *100`% of women in the study are carriers.  

```{r}
amyloid = 
  read_csv("data_mci/mci_amyloid.csv", skip = 1) |>
  janitor::clean_names() |>
  rename(id = study_id) |>
  pivot_longer(
    baseline:time_8,
    names_to = "year",
    names_prefix = "time_",
    values_to = "value"
  ) |>
  mutate(
    year = replace(year, year == "baseline", "0"))
```

Read the file first and use pivot_longer so that I could do the `anti_join` and `inner_join` to compare and summarize the datasets. `amyloid` datasets has `r nrow(amyloid)` participants and `r ncol(amyloid)` variables.

```{r}
only_data<-anti_join(baseline_mci,amyloid)
onlt_data_2<-anti_join(amyloid,baseline_mci)
```

After using `anti_join` function for both datasets, notice that `r nrow(only_data)`participants in the `baseline` are not included in `amyloid` while 2 participants in the `amyloid` are not included in the `baseline`.  

```{r}
total_data = inner_join(baseline_mci, amyloid, by="id")

write_csv(total_data, "~/Desktop/p8105_hw2_sc5352/data_mci/data_mci.csv")
```

After merging two datasets, there are `r nrow(total_data)` rows. Each row shows id, current age, sex, education, apoe4, and the value of mci during different time period.


